Structured Versus Decorated Cospans

### 2020-11-10 cospans.tex
worked on intro, changed all \mathsf{X} to \X

An `open system' is any sort of system that can interact with the outside world.  Open systems
are nicely modeled using cospans, which are diagrams of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the \define{apex} of the cospan, $a_1$ and $a_2$ its \define{feet}, and $i$ and $o$ its \define{legs}.   The feet of a cospan describe `interfaces' of the corresponding open system, and
open systems modeled as cospans can composed via pushout.  However, we often want the system itself---the apex---to have more structure than its interfaces.   This led Fong to develop a theory of `decorated' cospans \cite{Fong}.

@@

\begin{thm}\label{SC}
Given a category $\mathsf{X}$ with finite colimits and a category $\mathsf{A}$ with finite coproducts and a finite coproduct preserving functor $L \maps \mathsf{A} \to \mathsf{X}$ with $\mathsf{A}$ and $\mathsf{X}$, there exists a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ which has:
\begin{itemize}
\item{objects given by objects of $\mathsf{A}$,}
\item{vertical 1-morphisms given by morphisms of $\mathsf{A}$,}
\item{horizontal 1-cells $a\tickar b$ given by cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}}
\item 2-morphisms given by maps of cospans of $\mathsf{X}$ of the form:
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
\end{itemize}
Composition of horizontal 1-cells and 2-morphisms is given by pushouts in $\mathsf{X}$ and tensoring of objects is under binary coproducts in $\mathsf{A}$.
\end{thm}
\begin{proof}
See the first two authors' work on structured cospans \cite{BC2}.
\end{proof}

@@

which is a cospan in $\mathsf{X}$ of the form:

@@

where $R\maps \mathsf{X} \to \mathsf{A}$ is the right adjoint to the functor $L \maps \mathsf{A} \to \mathsf{X}$ and $\eta \maps 1_{\mathsf{A}} \to RL$ is the unit of the adjunction $L \dashv R$ which is an isomorphism since $L$ is fully faithful. \textbf{Hopefully...actually, it's a `lari'!} Similarly, the image of a 2-morphism $\alpha \maps M \to N$ in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$:

@@

together with a morphism $\iota \maps F(R(\alpha))(x) \to x^\prime$ in $F(R(x^\prime)) \subseteq \mathrm{X}$ which comes from the Grothendieck construction of the pseudofunctor $F \maps \mathsf{A} \to \Cat$.  That $\mathbb{E}_0$ is a functor is clear. For $\mathbb{E}_1$, given two vertically composable 2-morphisms $M$ and $M'$ in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$,

@@

To see that this functor is essentially surjective, given a horizontal 1-cell in $F\mathbb{C}\textnormal{sp}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$c_2$};
\node (D) at (3,0) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
we can find a 2-isomorphism in $F\mathbb{C}\textnormal{sp}$ whose codomain is the above horizontal 1-cell and whose domain is the image of the following horizontal 1-cell in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c_1)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
with the 2-isomorphism in $F\mathbb{C}\textnormal{sp}$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$R(x)$};
\node (C) at (2,0) {$c_2$};
\node (A') at (0,-1) {$c_1$};
\node (B') at (1,-1) {$c$};
\node (C') at (2,-1) {$c_2$};
\node (D) at (3,0) {$x \in F(R(x))$};
\node (D') at (3,-1) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(i^\prime) \eta_{c_1}$} (B)
(C) edge node[above]{$R(o^\prime) \eta_{c_2}$} (B)
(A') edge node[above]{$i$} (B')
(C') edge node[above]{$o$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{${(R(e) \eta_c)}^{-1}$} (B')
(C) edge node [right]{$1$} (C');
\end{tikzpicture}
\]
$$\iota \maps F({(R(e)\eta_c)}^{-1})(x) \to x$$
where $e \maps L(c) \to x$ is given by the map from the trivial decoration on $c$ to $x \in F(c)$. %The object and arrow components $\mathbb{E}_0$ and $\mathbb{E}_1$ satisfy the equations $S \mathbb{E}_1 = \mathbb{E}_0 S$ and $T \mathbb{E}_1 = \mathbb{E}_0 T$.

To show that the double functor $\mathbb{E}$ is fully faithful, we need to show that the map  $$\mathbb{E}_1 \maps _f { _L \mathbb{C}\textnormal{sp}(\mathsf{X})}_g(M,N) \to {_{\mathbb{E}(f)} {F\mathbb{C}\textnormal{sp}}_{\mathbb{E}(g)}}(\mathbb{E}(M),\mathbb{E}(N))$$ is bijective for arbitrary vertical 1-morphisms $f$ and $g$ and horizontal 1-cells $M$ and $N$ of $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$.  Consider a 2-morphism in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ with horizontal source and target $M$ and $N$, respectively and vertical source and target $f$ and $g$, respectively:

@@

Thus the set $$_f { _L \mathbb{C}\textnormal{sp}(\mathsf{X})}_g(M,N)$$ consists of triples $$(f,\alpha,g)$$ rendering the above diagram commutative where $f$ and $g$ are morphisms of $\mathsf{A}$ and $\alpha$ is a morphism of $\mathsf{X}$. The image of the above 2-morphism under the double functor $\mathbb{E}$ is given by:

@@

Thus the set $$_{\mathbb{E}(f)} {F\mathbb{C}\textnormal{sp}}_{\mathbb{E}(g)}(\mathbb{E}(M),\mathbb{E}(N))$$ consists of 4-tuples $$(f,R(\alpha),g,\iota)$$ rendering the above diagram commutative and where $f,g$ and $R(\alpha)$ are morphisms of $\mathsf{A}$ and $\iota$ is a morphism in $F(R(x'))$. The morphisms $R(\alpha) \maps R(x) \to R(x^\prime)$ and $\iota \maps F(R(\alpha))(x) \to x^\prime$ together determine the morphism $\alpha \maps x \to x^\prime$ in $\mathrm{X}$ and conversely; given two objects $x=(c,x \in F(c))$ and $x^\prime=(c^\prime,x^\prime \in F(c^\prime))$ of $\mathsf{X}=\inta{F}$, a morphism from $\alpha \maps x \to x^\prime$ is a pair $$(h \maps c \to c^\prime, \iota \maps F(h)(x) \to x^\prime)$$ where $h \maps c \to c^\prime$ is given by $R(\alpha) \maps R(x) \to R(x^\prime)$. This shows that $\mathbb{E}$ is fully faithful. \textbf{At least in my favorite example...}

@@

Next, given composable horizontal 1-cells $M$ and $N$ in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$:

@@

together with a morphism of decorations which is obtained as follows. First, as $R \maps \mathsf{X} \to \mathsf{A}$ preserves finite colimits, we have an isomorphism $$\kappa \maps R(x) +_{R(L(c_2))} R(x^\prime) \to R(x+_{L(c_2)}x^\prime).$$ Also, since the left adjoint $L \maps \mathsf{A} \to \mathsf{X}$ is fully faithful, the unit of the adjunction $L \dashv R$ at the object $c_2$ gives an isomorphism $\eta_{c_2} \maps c_2 \to R(L(c_2))$ which results in an isomorphism $$j_{\eta_{c_2}} \maps R(x) +_{c_2} R(x^\prime) \to R(x) +_{R(L(c_2))} R(x^\prime).$$ Composing these two results in an isomorphism $$\sigma \mapseqq \kappa j_{\eta_{c_2}} \maps R(x) +_{c_2} R(x^\prime) \to R(x+_{L(c_2)}x^\prime).$$

@@

Next, we show that if both double categories $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ and $F\mathbb{C}\textnormal{sp}$ are symmetric monoidal, as they are if both $\mathsf{A}$ and $\mathsf{X}$ have finite colimits, then this equivalence of double categories $\mathbb{E} \maps _L \mathbb{C}\textnormal{sp}(\mathsf{X}) \to F\mathbb{C}\textnormal{sp}$ will be symmetric monoidal. First, note that we have an isomorphism $\epsilon \maps 1_{F\mathbb{C}\textnormal{sp}} \to \mathbb{E}(1_{_L \mathbb{C}\textnormal{sp}(\mathsf{X})})$ and natural isomorphisms $\mu_{c_1,c_2} \maps \mathbb{E}(c_1) \otimes \mathbb{E}(c_2) \to \mathbb{E}(c_1 \otimes c_2)$ for every pair of objects $c_1,c_2 \in {_L \mathbb{C} \textnormal{sp}(\mathsf{X})}$ both of which are given by identities since both double categories $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ and $F\mathbb{C}\textnormal{sp}$ have $\mathsf{A}$ as their category of objects and $\mathbb{E}_0=\id_{\mathsf{A}}$. The diagrams utilizing these maps that are required to commute do so trivially.

For the arrow component $\mathbb{E}_1$, we have an isomorphism $\delta \maps U_{1_{F\mathbb{C}\textnormal{sp}}} \to \mathbb{E}(U_{1_{_L \mathbb{C}\textnormal{sp}(\mathsf{X})}})$ where the horizontal 1-cell $U_{1_{F\mathbb{C}\textnormal{sp}}}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1,0) {$1_\mathsf{A}$};
\node (C) at (2,0) {$1_\mathsf{A}$};
\node (D) at (3,0) {$!_{1_{\mathsf{A}}} \in F(1_\mathsf{A})$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B);
\end{tikzpicture}
\]
where $!_{1_{\mathsf{A}}} = \phi \maps 1 \to F(1_\mathsf{A})$ is the trivial decoration which comes from the structure of the symmetric lax monoidal pseudofunctor $F \maps \mathsf{A} \to \Cat$. The horizontal 1-cell $U_{1_{_L \mathbb{C}\textnormal{sp}(\mathsf{X})}}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(1_\mathsf{A})$};
\node (B) at (1,0) {$L(1_\mathsf{A})$};
\node (C) at (2,0) {$L(1_\mathsf{A})$};
%\node (D) at (3,0.5) {$I \in F(1_\mathrm{A})$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B);
\end{tikzpicture}
\]
where here we make use of the fact that the left adjoint $L \maps (\mathsf{A},+,1_\mathsf{A}) \to (\mathsf{X},+,1_\mathsf{X})$ preserves all colimits and thus $L(1_\mathsf{A}) \cong 1_\mathsf{X}$. The horizontal 1-cell $\mathbb{E}(U_{1_{_L \mathbb{C} \textnormal{sp}(\mathsf{X})}})$ is then given by the pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1.25,0) {$R(L(1_\mathsf{A}))$};
\node (C) at (2.5,0) {$1_\mathsf{A}$};
\node (D) at (4.5,0) {$!_{R(L(1_\mathsf{A}))} \in F(R(L(1_\mathsf{A})))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{${\eta_{1_\mathsf{A}}}$} (B)
(C) edge node[above]{${\eta_{1_\mathsf{A}}}$} (B);
\end{tikzpicture}
\]
The isomorphism $\delta$ is then given by the 2-morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1,0) {$1_{\mathsf{A}}$};
\node (C) at (2,0) {$1_\mathsf{A}$};
\node (A') at (0,-1) {$1_\mathsf{A}$};
\node (B') at (1,-1) {$R(L(1_\mathsf{A}))$};
\node (C') at (2,-1) {$1_\mathsf{A}$};
\node (D) at (3,0) {$!_{1_\mathsf{A}} \in F(1_\mathsf{A})$};
\node (D') at (3.75,-1) {$!_{R(L(1_\mathsf{A}))} \in F(R(L(1_\mathsf{A})))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B)
(A') edge node[above]{${\eta_{1_\mathsf{A}}}$} (B')
(C') edge node[above]{${\eta_{1_\mathsf{A}}}$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{${\eta_{1_\mathsf{A}}}$} (B')
(C) edge node [left]{$1$} (C');
\end{tikzpicture}
\]
$$\iota_{{\eta_{1_\mathsf{A}}}} \maps F({\eta_{1_\mathsf{A}}})(!_{1_\mathsf{A}})) \to !_{R(L(1_\mathsf{A}))}$$
of $F\mathbb{C}\textnormal{sp}$. This is just the natural isomorphism $\mathbb{E}_{1_\mathsf{A}}$ from earlier.

Given two horizontal 1-cells $M$ and $N$ of $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$:

@@

where $\kappa$ is the isomorphism which comes from $R \maps \mathsf{X} \to \mathsf{A}$ preserving finite colimits.

@@

To see this, let $M_1,M_2$ and $M_3$ be horizontal 1-cells in $_L \mathbb{C}\textnormal{sp}(\mathsf{X})$ given by:

@@

\node (B) at (3,0) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\mathsf{X})}) \otimes \mathbb{E}(M_1)$};

@@

\node (D) at (3,-1) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\mathsf{X})} \otimes M_1)$};

@@

			\node (d) at (7.5,0) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\mathsf{X})}) \otimes \mathbb{E}(M_1)$};

@@

			\node (d4) at (7.5,-1) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\mathsf{X})} \otimes M_1)$};

@@

                                (b5) edge node[left]{$R(\lambda_{\mathsf{X}})$} (b6)

@@

\node (C) at (4,0) {$F(R(\lambda_{\mathsf{X}})(\mu_{L(1_\mathsf{A}),x_1}))(!_{R(L(1_\mathsf{A}))} +x_1$)};

@@

\node (E) at (4,-1) {$F(R(\lambda_\mathsf{X}))(x_{!+1})$};
(A) edge node[above]{$F(R(\lambda_{\mathsf{X}})(\mu_{L(1_\mathsf{A}),x_1}))(\iota_2)$} (C)

@@

(C) edge node[right]{$F(R(\lambda_\mathsf{X}))(\iota_3)$} (E);

>>>>

An `open system' is any sort of system that can interact with the outside world.  Open systems
are nicely modeled using cospans, which are diagrams of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the \define{apex}, $a_1$ and $a_2$ the \define{feet}, and $i$ and $o$ the \define{legs}.   The feet describe `interfaces' of the corresponding open system.  However, we often want the system itself---the apex---to have more structure than its interfaces.   This led Fong to develop a theory of `decorated' cospans \cite{Fong}.

@@

\begin{thm}\label{SC}
Given a category $\X$ with finite colimits and a category $\mathsf{A}$ with finite coproducts and a finite coproduct preserving functor $L \maps \mathsf{A} \to \X$ with $\mathsf{A}$ and $\X$, there exists a symmetric monoidal double category $_L \mathbb{C}\textnormal{sp}(\X)$ which has:
\begin{itemize}
\item{objects given by objects of $\mathsf{A}$,}
\item{vertical 1-morphisms given by morphisms of $\mathsf{A}$,}
\item{horizontal 1-cells $a\tickar b$ given by cospans of $\X$ of the form:
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}}
\item 2-morphisms given by maps of cospans of $\X$ of the form:
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
\end{itemize}
Composition of horizontal 1-cells and 2-morphisms is given by pushouts in $\X$ and tensoring of objects is under binary coproducts in $\mathsf{A}$.
\end{thm}
\begin{proof}
See the first two authors' work on structured cospans \cite{BC2}.
\end{proof}

@@

which is a cospan in $\X$ of the form:

@@

where $R\maps \X \to \mathsf{A}$ is the right adjoint to the functor $L \maps \mathsf{A} \to \X$ and $\eta \maps 1_{\mathsf{A}} \to RL$ is the unit of the adjunction $L \dashv R$ which is an isomorphism since $L$ is fully faithful. \textbf{Hopefully...actually, it's a `lari'!} Similarly, the image of a 2-morphism $\alpha \maps M \to N$ in $_L \mathbb{C}\textnormal{sp}(\X)$:

@@

together with a morphism $\iota \maps F(R(\alpha))(x) \to x^\prime$ in $F(R(x^\prime)) \subseteq \mathrm{X}$ which comes from the Grothendieck construction of the pseudofunctor $F \maps \mathsf{A} \to \Cat$.  That $\mathbb{E}_0$ is a functor is clear. For $\mathbb{E}_1$, given two vertically composable 2-morphisms $M$ and $M'$ in $_L \mathbb{C}\textnormal{sp}(\X)$,

@@

To see that this functor is essentially surjective, given a horizontal 1-cell in $F\mathbb{C}\textnormal{sp}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$c_2$};
\node (D) at (3,0) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
we can find a 2-isomorphism in $F\mathbb{C}\textnormal{sp}$ whose codomain is the above horizontal 1-cell and whose domain is the image of the following horizontal 1-cell in $_L \mathbb{C}\textnormal{sp}(\X)$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(c_1)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(c_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
with the 2-isomorphism in $F\mathbb{C}\textnormal{sp}$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c_1$};
\node (B) at (1,0) {$R(x)$};
\node (C) at (2,0) {$c_2$};
\node (A') at (0,-1) {$c_1$};
\node (B') at (1,-1) {$c$};
\node (C') at (2,-1) {$c_2$};
\node (D) at (3,0) {$x \in F(R(x))$};
\node (D') at (3,-1) {$x \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$R(i^\prime) \eta_{c_1}$} (B)
(C) edge node[above]{$R(o^\prime) \eta_{c_2}$} (B)
(A') edge node[above]{$i$} (B')
(C') edge node[above]{$o$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{${(R(e) \eta_c)}^{-1}$} (B')
(C) edge node [right]{$1$} (C');
\end{tikzpicture}
\]
$$\iota \maps F({(R(e)\eta_c)}^{-1})(x) \to x$$
where $e \maps L(c) \to x$ is given by the map from the trivial decoration on $c$ to $x \in F(c)$. %The object and arrow components $\mathbb{E}_0$ and $\mathbb{E}_1$ satisfy the equations $S \mathbb{E}_1 = \mathbb{E}_0 S$ and $T \mathbb{E}_1 = \mathbb{E}_0 T$.

To show that the double functor $\mathbb{E}$ is fully faithful, we need to show that the map  $$\mathbb{E}_1 \maps _f { _L \mathbb{C}\textnormal{sp}(\X)}_g(M,N) \to {_{\mathbb{E}(f)} {F\mathbb{C}\textnormal{sp}}_{\mathbb{E}(g)}}(\mathbb{E}(M),\mathbb{E}(N))$$ is bijective for arbitrary vertical 1-morphisms $f$ and $g$ and horizontal 1-cells $M$ and $N$ of $_L \mathbb{C}\textnormal{sp}(\X)$.  Consider a 2-morphism in $_L \mathbb{C}\textnormal{sp}(\X)$ with horizontal source and target $M$ and $N$, respectively and vertical source and target $f$ and $g$, respectively:

@@

Thus the set $$_f { _L \mathbb{C}\textnormal{sp}(\X)}_g(M,N)$$ consists of triples $$(f,\alpha,g)$$ rendering the above diagram commutative where $f$ and $g$ are morphisms of $\mathsf{A}$ and $\alpha$ is a morphism of $\X$. The image of the above 2-morphism under the double functor $\mathbb{E}$ is given by:

@@

Thus the set $$_{\mathbb{E}(f)} {F\mathbb{C}\textnormal{sp}}_{\mathbb{E}(g)}(\mathbb{E}(M),\mathbb{E}(N))$$ consists of 4-tuples $$(f,R(\alpha),g,\iota)$$ rendering the above diagram commutative and where $f,g$ and $R(\alpha)$ are morphisms of $\mathsf{A}$ and $\iota$ is a morphism in $F(R(x'))$. The morphisms $R(\alpha) \maps R(x) \to R(x^\prime)$ and $\iota \maps F(R(\alpha))(x) \to x^\prime$ together determine the morphism $\alpha \maps x \to x^\prime$ in $\mathrm{X}$ and conversely; given two objects $x=(c,x \in F(c))$ and $x^\prime=(c^\prime,x^\prime \in F(c^\prime))$ of $\X=\inta{F}$, a morphism from $\alpha \maps x \to x^\prime$ is a pair $$(h \maps c \to c^\prime, \iota \maps F(h)(x) \to x^\prime)$$ where $h \maps c \to c^\prime$ is given by $R(\alpha) \maps R(x) \to R(x^\prime)$. This shows that $\mathbb{E}$ is fully faithful. \textbf{At least in my favorite example...}

@@

Next, given composable horizontal 1-cells $M$ and $N$ in $_L \mathbb{C}\textnormal{sp}(\X)$:

@@

together with a morphism of decorations which is obtained as follows. First, as $R \maps \X \to \mathsf{A}$ preserves finite colimits, we have an isomorphism $$\kappa \maps R(x) +_{R(L(c_2))} R(x^\prime) \to R(x+_{L(c_2)}x^\prime).$$ Also, since the left adjoint $L \maps \mathsf{A} \to \X$ is fully faithful, the unit of the adjunction $L \dashv R$ at the object $c_2$ gives an isomorphism $\eta_{c_2} \maps c_2 \to R(L(c_2))$ which results in an isomorphism $$j_{\eta_{c_2}} \maps R(x) +_{c_2} R(x^\prime) \to R(x) +_{R(L(c_2))} R(x^\prime).$$ Composing these two results in an isomorphism $$\sigma \mapseqq \kappa j_{\eta_{c_2}} \maps R(x) +_{c_2} R(x^\prime) \to R(x+_{L(c_2)}x^\prime).$$

@@

Next, we show that if both double categories $_L \mathbb{C}\textnormal{sp}(\X)$ and $F\mathbb{C}\textnormal{sp}$ are symmetric monoidal, as they are if both $\mathsf{A}$ and $\X$ have finite colimits, then this equivalence of double categories $\mathbb{E} \maps _L \mathbb{C}\textnormal{sp}(\X) \to F\mathbb{C}\textnormal{sp}$ will be symmetric monoidal. First, note that we have an isomorphism $\epsilon \maps 1_{F\mathbb{C}\textnormal{sp}} \to \mathbb{E}(1_{_L \mathbb{C}\textnormal{sp}(\X)})$ and natural isomorphisms $\mu_{c_1,c_2} \maps \mathbb{E}(c_1) \otimes \mathbb{E}(c_2) \to \mathbb{E}(c_1 \otimes c_2)$ for every pair of objects $c_1,c_2 \in {_L \mathbb{C} \textnormal{sp}(\X)}$ both of which are given by identities since both double categories $_L \mathbb{C}\textnormal{sp}(\X)$ and $F\mathbb{C}\textnormal{sp}$ have $\mathsf{A}$ as their category of objects and $\mathbb{E}_0=\id_{\mathsf{A}}$. The diagrams utilizing these maps that are required to commute do so trivially.

For the arrow component $\mathbb{E}_1$, we have an isomorphism $\delta \maps U_{1_{F\mathbb{C}\textnormal{sp}}} \to \mathbb{E}(U_{1_{_L \mathbb{C}\textnormal{sp}(\X)}})$ where the horizontal 1-cell $U_{1_{F\mathbb{C}\textnormal{sp}}}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1,0) {$1_\mathsf{A}$};
\node (C) at (2,0) {$1_\mathsf{A}$};
\node (D) at (3,0) {$!_{1_{\mathsf{A}}} \in F(1_\mathsf{A})$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B);
\end{tikzpicture}
\]
where $!_{1_{\mathsf{A}}} = \phi \maps 1 \to F(1_\mathsf{A})$ is the trivial decoration which comes from the structure of the symmetric lax monoidal pseudofunctor $F \maps \mathsf{A} \to \Cat$. The horizontal 1-cell $U_{1_{_L \mathbb{C}\textnormal{sp}(\X)}}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(1_\mathsf{A})$};
\node (B) at (1,0) {$L(1_\mathsf{A})$};
\node (C) at (2,0) {$L(1_\mathsf{A})$};
%\node (D) at (3,0.5) {$I \in F(1_\mathrm{A})$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B);
\end{tikzpicture}
\]
where here we make use of the fact that the left adjoint $L \maps (\mathsf{A},+,1_\mathsf{A}) \to (\X,+,1_\X)$ preserves all colimits and thus $L(1_\mathsf{A}) \cong 1_\X$. The horizontal 1-cell $\mathbb{E}(U_{1_{_L \mathbb{C} \textnormal{sp}(\X)}})$ is then given by the pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1.25,0) {$R(L(1_\mathsf{A}))$};
\node (C) at (2.5,0) {$1_\mathsf{A}$};
\node (D) at (4.5,0) {$!_{R(L(1_\mathsf{A}))} \in F(R(L(1_\mathsf{A})))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{${\eta_{1_\mathsf{A}}}$} (B)
(C) edge node[above]{${\eta_{1_\mathsf{A}}}$} (B);
\end{tikzpicture}
\]
The isomorphism $\delta$ is then given by the 2-morphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$1_\mathsf{A}$};
\node (B) at (1,0) {$1_{\mathsf{A}}$};
\node (C) at (2,0) {$1_\mathsf{A}$};
\node (A') at (0,-1) {$1_\mathsf{A}$};
\node (B') at (1,-1) {$R(L(1_\mathsf{A}))$};
\node (C') at (2,-1) {$1_\mathsf{A}$};
\node (D) at (3,0) {$!_{1_\mathsf{A}} \in F(1_\mathsf{A})$};
\node (D') at (3.75,-1) {$!_{R(L(1_\mathsf{A}))} \in F(R(L(1_\mathsf{A})))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$1$} (B)
(A') edge node[above]{${\eta_{1_\mathsf{A}}}$} (B')
(C') edge node[above]{${\eta_{1_\mathsf{A}}}$} (B')
(A) edge node [left]{$1$} (A')
(B) edge node [left]{${\eta_{1_\mathsf{A}}}$} (B')
(C) edge node [left]{$1$} (C');
\end{tikzpicture}
\]
$$\iota_{{\eta_{1_\mathsf{A}}}} \maps F({\eta_{1_\mathsf{A}}})(!_{1_\mathsf{A}})) \to !_{R(L(1_\mathsf{A}))}$$
of $F\mathbb{C}\textnormal{sp}$. This is just the natural isomorphism $\mathbb{E}_{1_\mathsf{A}}$ from earlier.

Given two horizontal 1-cells $M$ and $N$ of $_L \mathbb{C}\textnormal{sp}(\X)$:

@@

where $\kappa$ is the isomorphism which comes from $R \maps \X \to \mathsf{A}$ preserving finite colimits.

@@

To see this, let $M_1,M_2$ and $M_3$ be horizontal 1-cells in $_L \mathbb{C}\textnormal{sp}(\X)$ given by:

@@

\node (B) at (3,0) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\X)}) \otimes \mathbb{E}(M_1)$};

@@

\node (D) at (3,-1) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\X)} \otimes M_1)$};

@@

			\node (d) at (7.5,0) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\X)}) \otimes \mathbb{E}(M_1)$};

@@

			\node (d4) at (7.5,-1) {$\mathbb{E}(1_{ _L \mathbb{C}\textnormal{sp}(\X)} \otimes M_1)$};

@@

                                (b5) edge node[left]{$R(\lambda_{\X})$} (b6)

@@

\node (C) at (4,0) {$F(R(\lambda_{\X})(\mu_{L(1_\mathsf{A}),x_1}))(!_{R(L(1_\mathsf{A}))} +x_1$)};

@@

\node (E) at (4,-1) {$F(R(\lambda_\X))(x_{!+1})$};
(A) edge node[above]{$F(R(\lambda_{\X})(\mu_{L(1_\mathsf{A}),x_1}))(\iota_2)$} (C)

@@

(C) edge node[right]{$F(R(\lambda_\X))(\iota_3)$} (E);
where $x_{!+1}$ is the decoration $x_1$ on the object $R(L(1_\mathsf{A})+x_1) \in \mathsf{A}$. The above square commutes because $$F(\lambda_\mathsf{A})(!_{1_\mathsf{A}}+x_1) = F(R(\lambda_\X)(\mu_{L(1_\mathsf{A}),x_1})(\eta_{1_\mathsf{A}}+1))(!_{1_\mathsf{A}}+x_1)$$ as the corresponding left unitality square for the finite colimit preserving functor $R \maps (\X,1_\X,+) \to (\mathsf{A},1_\mathsf{A},+)$ commutes. The right unitality square is similar. The associator hexagon:
